Cultural evolutionary dynamics with prestige bias

### 2022-11-08 ms/ms.tex
update

\makeatletter

\def\blfootnote{\gdef\@thefnmark{}\@footnotetext}

\makeatother

@@

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small,labelfont=bf]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage{varioref}
\usepackage[unicode=true,colorlinks=true,urlcolor=blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
\usepackage{cjhebrew}

% unbreakable dashes
\usepackage[shortcuts]{extdash}
% footnotes
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

% Supplementary
\newcommand{\beginsupplement}{%
      	\setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

\begin{document}

\begin{titlepage}
	\centering
	\includegraphics[width=0.4\textwidth]{ReichmanLogo.png}\par\vspace{1cm}
	{\huge Reichman University \par}
	{\Large Efi Arazi School of Computer Science \par}
	{\Large M.Sc. Program - Research Track \par}

	\vspace{1.5cm}
	{\Huge Cultural Evolutionary Dynamics with Success and Influence Bias\par}
	\vspace{1.5cm}
	{\large by\par}
	{\Large\bfseries Saar Egozi\par}
	\vspace{1cm}
	%{\large Supervised by Dr. Yoav Ram\par}
	%{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv 6997801, Israel\par}
%	{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv 6997801, Israel \par}

	\vspace{1cm}
	{M.Sc. dissertation, submitted in partial fulfillment of the requirements\par}
	{for the M.Sc. degree, research track, School of Computer Science\par}
	{Reichman University}

	\vspace{1cm}
	% Bottom of the page
	{\large \today \par}

\end{titlepage}

% title page stuff
%\title{Cultural Evolutionary Dynamics with Success and Influence biases}

%\renewcommand\Affilfont{\small}
%\author[1]{Saar Egozi}
%\author[2,3,*]{Yoav Ram}
%\affil[1]{School of Computer Science, Reichman University, Herzliya 4610101, Israel}
%\affil[2]{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv 6997801, Israel}
%\affil[3]{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv 6997801, Israel}
%\affil[*]{Corresponding author: yoav@yoavram.com, ORCID 0000-0002-9653-4458}

%\date{\today}

%\maketitle

This work was carried out under the supervision of Dr. Yoav Ram as part of the M.Sc.
program of Efi Arazi School of Computer Science, Reichman University.

@@

\section{Abstract}
A common bias in cultural transmission is to copy whoever appears more successful. This ``success bias'' only depends on the performance of the role-model.
Here, we propose an additional transmission bias that may be common in cultural transmission: ``influence bias'', in which the choice of a role-model depends on the number of individuals that have already copied from that role-model.
We combine success and influence bias together to form a ``prestige bias'', and analyze its effects on cultural-evolutionary dynamics using mathematical analysis and stochastic simulations.
We find analytic approximations to the stochastic role-model choice process, which facilitate further mathematical analysis and reduces the computational complexity of simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
We also find approximations to the fixation probability and the fixation time of an invading cultural trait, in both constant and changing environments. 
These approximations are similar to Kimura's approximations for population-genetic models.
These similarities show that success bias effectively plays the role of natural selection, whereas influence bias effectively plays the role of genetic drift.

Influence bias also accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model extends the literature that attempt to describe of cultural transmission, 
especially in human societies where social media is popular and thus influence bias may be expected. 
Further work is needed to test if this model could predict various phenomena in cultural evolution.

\blfootnote{This work is in preparation for submission to a high-impact journal.}
\pagebreak

\tableofcontents

\section{Introduction}

@@

\subsection{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, we assume homogeneous $\alpha$ for simplicity and for easier mathematical analysis.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + C\alpha'\beta(\hat{A})} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + \alpha'C}
\end{equation}
where $C$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model and $\alpha'$ is the odd ratio of the bias weight $\alpha'= \frac{\alpha}{1-\alpha}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\subsection{Approximations}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}. Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two such approximations, summarized here and explained in detail below. In both we assume either $\alpha_j$ is a trait of role-model $j$ only (contrary to $\alpha_{ij}$ that included the relation with copier $i$), or a completely homogeneous $\alpha$.
\begin{enumerate}
\item 
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,
\begin{enumerate}\item the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$). This means that there's no actual estimation error in the population, just a fixed bias in every copier estimation.
\item the expected number of copiers of each role-model equals its relative biased trait value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the bias weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated trait value. 
\end{enumerate}
\item The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$), meaning there's no meaning for the order of copiers. 
Therefore, the number of copiers of all role-models, $\vec{K}_i = (K_{i,1}, \ldots, K_{i,N})$, follows a Dirichlet-Multinomial distribution,
$\vec{K}_i \sim \mathit{DM}\big(N, \vec{G}_1\big)$, where $\vec{G}_1 = (G_{1,1}, \ldots, G_{1,N})$ is the prestige vector of all role-models in the "eyes" of the first copier.
Note that here $G_{i,j}$ is only a function of the trait values $A_j$ and the bias weights $\alpha_j$, as the estimation error is uniform, meaning all copiers are identical.
\end{enumerate}

\subsection{Generalized binomial distribution}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is parameterized by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).
\\

\begin{result}[GBD approximation]\label{res:GBD}
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$. 
\end{result}

\begin{proof}
We denote $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ as the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ our of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1),
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) by \citet{GBD}, which completes the proof.
\end{proof}

@@

\begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{../figures/final/GBD_validation.pdf}
  \caption{
  \textbf{Numerical validation of the GBD approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models.
  Here, population size, $N=2,000$; bias weight, $\alpha=0.1$; idea phenotype value, $\hat{A}=1$; role-model traits $\vec{A} \sim N(0,1)$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:GBD}
\end{figure}

Although basic, \cref{fig:GBD} shows good fit of the GBD approximation.
This validation is initial, and the more extensive validations we do on the Dirichlet-Multinomial approximation, because it is what we will use in our analysts.

\subsection{Dirichlet-Multinomial distribution}

@@

\begin{figure}[H]
    \includegraphics[width=\linewidth]{../figures/final/choose_bias.pdf}
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{A}) and estimation error (\textbf{B}) decrease during the role-model choosing process, demonstrating that influence becomes more favored as more copiers have made their choice.
However, when $\alpha$ is homogeneous (\textbf{C}), the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\section{Discussion}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be copied more often due to transmission biases. 
A common bias is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be accurately estimated.
However, it has been suggested that because it is hard to estimate success, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful in other traits \citep{fijian_social_bias}.

@@

{\small
\section{Acknowledgements}
We thank Marcus Feldman, Martin Pontz, and Tal Simon for discussions and comments.
This work was supported in part by 
the Israel Science Foundation (YR 552/19),
Minerva Stiftung Center for Lab Evolution (YR),
and 
the John Templeton Foundation (YR 61809).
}

\section{References}
\nolinenumbers
\bibliography{refs}
%\bibliographystyle{agsm}
\bibliographystyle{apalike}

@@

\appendix
\section{General Binomial Distribution approximation proof}\label{append:GBD}

@@

\section{Drift and diffusion in a constant environment proof}\label{append:drift_diff_const}

@@

\section{Drift and diffusion in a changing environment proof}\label{append:drift_diff_chang}
\paragraph{Proving drift and diffusion terms in a changing environment.}
Let $x$ be the initial frequency of the invading phenotype and $X_t$ is the number of individuals with the phenotype at time $t$.
Then,
\begin{equation*}
E[X_t/N - x] \simeq x(1-x) S_t / N_e \;, 
\quad
\text{and}
\quad
V(X_t/N-x) \simeq  t x(1-x) / N_e \;,
\end{equation*}
where $S_t=\sum\limits_{i=1}^{t} N (1-\beta_t)$.

@@

Using \cref{eq:const_var} we get
\begin{equation}
V\left(\frac{X_{t+1}}{N} \bigg|X_t \right) 
= \frac{1}{N_e}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \;,
\end{equation}
and using the equation $y'(1-y') \simeq y(1-y)$ on the first part of \cref{eq:ch_var} we get
\begin{equation}\label{eq:ch_var2}
E\left[V\left(\frac{X_{t+1}}{N} \bigg|X_t \right)\right] 
= \frac{1}{N_e}E\left[\frac{X_t}{N}\left(1- \frac{X_t}{N}\right) \right] \simeq \frac{1}{N_e} x(1-x) \;.
\end{equation}
Moving on to simplify the second part of \cref{eq:ch_var} using \cref{eq:ch_var1},
\begin{equation}
V\left(E\left[\frac{X_{t+1}}{N} \bigg|X_t \right]\right) 
= V\left(\frac{X_t}{N} + \frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \;.\right) \right)
\end{equation}
Now, because $\frac{X_t}{N}$ is a frequency, i.e $0 \leq X_t/N \leq 1$, we know that $V\left(\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \right)\leq\frac{1}{4}$. We therefore find that
\begin{equation}
V\left(\frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \right)
\leq \frac{1}{4N^2}s^2_{t+1} ;,
\end{equation}
and so it can be ignored.
Combining our equations we get
\begin{equation}
V\left(E\left[\frac{X_{t+1}}{N} \bigg|X_t \right]\right) 
= V\left(\frac{X_t}{N}\right) + O\left(\frac{1}{N^2}\right)\simeq V\left(\frac{X_t}{N}\right) \;.
\end{equation}
Using the induction assumption and \cref{eq:ch_var2},
\begin{equation}
V\left(\frac{X_{t+1}}{N}\right) 
\simeq \frac{1}{N_e}x(1-x) + \frac{1}{N_e}tx(1-x) \simeq \frac{1}{N_e}x(1-x)(t+1) \,
\end{equation}
which proves the diffusion term.
\end{proof}

>>>>

@@

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small,labelfont=bf]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage{varioref}
\usepackage[unicode=true,colorlinks=true,urlcolor=blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}
\usepackage[utf8x]{inputenc}
\usepackage[english]{babel}
% appendix
\usepackage[title,page]{appendix}
\usepackage{chngcntr}
% unbreakable dashes
\usepackage[shortcuts]{extdash}
% footnotes
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

% Supplementary
\newcommand{\beginsupplement}{%
      	\setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
        \setcounter{equation}{0}
        \renewcommand{\theequation}{A\arabic{equation}}%
}

@@

% line numbers
\usepackage[displaymath, mathlines]{lineno}
\renewcommand\linenumberfont{\normalfont\small\sffamily}
\linenumbers
\modulolinenumbers[2]

% Title page
\title{Cultural evolutionary dynamics with prestige bias}
% Authors
\renewcommand\Affilfont{\small}

\author[1]{Saar Egozi}
\author[2,3,$\dagger$]{Yoav Ram}

\affil[1]{School of Computer Science, Reichman University, Herzliya, Israel}
\affil[3]{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv, Israel}
\affil[3]{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv, Israel}
%\affil[*]{These authors contributed equally to this work}
\affil[$\dagger$]{Corresponding author: yoav@yoavram.com}

\begin{document}
\maketitle

\begin{abstract}
A common bias in cultural transmission is to copy whoever appears more successful. This ``success bias'' only depends on the performance of the role-model.
Here, we propose an additional transmission bias that may be common in cultural transmission: ``influence bias'', in which the choice of a role-model depends on the number of individuals that have already copied from that role-model.
We combine success and influence bias together to form a ``prestige bias'', and analyze its effects on cultural-evolutionary dynamics using mathematical analysis and stochastic simulations.
We find analytic approximations to the stochastic role-model choice process, which facilitate further mathematical analysis and reduces the computational complexity of simulations. We validate these approximations using simulations, and demonstrate their robustness to model assumptions.
We also find approximations to the fixation probability and the fixation time of an invading cultural trait, in both constant and changing environments. 
These approximations are similar to Kimura's approximations for population-genetic models.
These similarities show that success bias effectively plays the role of natural selection, whereas influence bias effectively plays the role of genetic drift.

Influence bias also accelerates the evolutionary dynamics, as can be expected in a \textit{rich-getting-richer} process.
Our model extends the literature that attempt to describe of cultural transmission, 
especially in human societies where social media is popular and thus influence bias may be expected. 
Further work is needed to test if this model could predict various phenomena in cultural evolution.
\end{abstract}

@@

\section*{Introduction}

@@

\subsection*{Dichotomous trait}
To allow for mathematical analysis of the model, we introduce a simplified version where the trait only has two phenotypes: the optimal phenotype $\hat{A}$ and the sub-optimal phenotype $A$. 
All role-models with the same phenotype will contribute to the probability of that phenotype to be transmitted, and thus influence is determined by the number of copiers that have already chosen a role-model with either phenotype.
In addition, we assume homogeneous $\alpha$ for simplicity and for easier mathematical analysis.
Therefore, the probability of the $i$-th copier to copy phenotype $A$ is
\begin{equation}\label{eq:binary-model}
G_{i,A} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + C\alpha'\beta(\hat{A})} = \frac{(N-C)\alpha'\beta(A) + K_{i,A}}{i-1 + (N-C)\alpha'\beta(A) + \alpha'C}
\end{equation}
where $C$ is the number of role-models with trait $\hat{A}$ and $K_{i,A}$ is the number of copiers that already chose $A$ when copier $i$ chooses a role-model and $\alpha'$ is the odd ratio of the bias weight $\alpha'= \frac{\alpha}{1-\alpha}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}.
Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two such approximations, summarized here and explained in detail below. In both we assume either $\alpha_j$ is a trait of role-model $j$ only (contrary to $\alpha_{ij}$ that included the relation with copier $i$), or a completely homogeneous $\alpha$.

\paragraph{Generalized binomial approximation.}  
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,

(i) the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$); and
(ii) the expected number of copiers of each role-model equals its relative biased trait value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the bias weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated trait value.

\paragraph{Dirichlet-Multinomial approximation.}
The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$), meaning there's no meaning for the order of copiers.

Note that here $G_{i,j}$ is only a function of the trait values $A_j$ and the bias weights $\alpha_j$, as the estimation error is uniform, meaning all copiers are identical.

\subsection*{Generalized binomial distribution}
The generalized binomial distribution (GBD) emerges from a series of dependent Bernoulli trials (in contrast to the standard binomial distribution in which trials are independent) and is parameterized by $GBD(n, p, \theta)$ where $n$ is the number of trials, $p$ is the probability of success of the first trial, and $\theta$ is the correlation between trials ($\theta=0$ gives the standard binomial distribution).
\\

\begin{result}[Generalized binomial approximation]\label{res:GBD}
The number of copiers of role-model $j$ after $i$ copiers have chosen a role-model follows the GBD, $K_{i,j} \sim GBD(i,\alpha_i\cdot\beta(A_j+e))$ if $e_i=e$ for all role-models $i=1, \ldots, N$. 
\end{result}

\begin{proof}
We denote $Q_j(k,i)=P(K_{i,j} = k \mid K_{i-1,j})$ as the probability that exactly $k$ out of $i$ copiers choose role-model $j$ given $K_{i-1,j}$ our of $i-1$ copiers chose role-model $j$. 
Using conditional probability and \cref{eq:recurrence},
\begin{equation}\label{recursive}
Q_j(k,i) = P_j(S_{i,j}=1 | k-1,i-1) \cdot Q_j(k-1,i-1) + P_j(S_{i,j} =0 | k,i-1) \cdot Q_j(k,i-1),
\end{equation}
where $S_{i,j} =1 $ when the $i$-th copier chooses role-model $j$.
\Cref{recursive} is equivalent to eq.~(2.1) by \citet{GBD}, which completes the proof.
\end{proof}
This result gives the following two corollaries on the expected number of followers of a given role-model by the end of the role-model choice process. The proof of \cref{cor:GBD} is in Appendix~\ref{sec:GBD}.
\\

@@

\begin{figure}[h]
    \includegraphics[width=0.7\linewidth]{../figures/final/GBD_validation.pdf}
  \caption{
  \textbf{Numerical validation of the GB approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations for both models.
  Here, population size, $N=2,000$; bias weight, $\alpha=0.1$; idea phenotype value, $\hat{A}=1$; role-model traits $\vec{A} \sim N(0,1)$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:GBD}
\end{figure}

Although basic, \cref{fig:GBD} shows good fit of the GB approximation.
This validation is initial, and the more extensive validations we do on the Dirichlet-Multinomial approximation, because it is what we will use in our analysts.

\subsection*{Dirichlet-Multinomial distribution approximation}

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/final/choose_bias.pdf}
  \caption{
  \textbf{Advantage of an adaptive success-bias weight.}
  Both success-bias weight $\alpha$ (\textbf{A}) and estimation error (\textbf{B}) decrease during the role-model choosing process, demonstrating that influence becomes more favored as more copiers have made their choice.
However, when $\alpha$ is homogeneous (\textbf{C}), the mean estimation error doesn't decrease, regardless of $\alpha$ or $\eta$.
The mean estimation error in the homogeneous $\alpha$ model is larger by a factor of $10$ than the adaptive $\alpha$ model.
Here, population size $N=200$; estimation error is normally distributed $e \sim N(0,\eta^2)$ with standard deviation $\eta=$0.0001 (blue), 0.001 (orange), 0.01 (green), plots are average of $300$ simulations.}	
  \label{fig:influence_advantage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Discussion}
During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
Some cultural traits or cultural role-models may be copied more often due to transmission biases. 
A common bias is success bias, in which copiers are more likely to copy a successful role-model. Many models assume that success can be accurately estimated.
However, it has been suggested that because it is hard to estimate success, a more common bias is \textit{prestige bias}---a bias towards role-models perceived to be successful in other traits \citep{fijian_social_bias}.

@@

{\small
\section*{Acknowledgements}
We thank Marcus Feldman, Martin Pontz, and Tal Simon for discussions and comments.
This work was supported in part by 
the Israel Science Foundation (YR 552/19),
Minerva Stiftung Center for Lab Evolution (YR),
and 
the John Templeton Foundation (YR 61809).
}

\section*{References}
\nolinenumbers
\bibliography{refs}
%\bibliographystyle{agsm}
\bibliographystyle{apalike}

@@

\begin{appendices}
\renewcommand{\theequation}{\thesection\arabic{equation}}
\counterwithin*{equation}{section}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{General binomial distribution approximation} \label{sec:GBD}

@@

\section{Drift and diffusion in a constant environment} \label{sec:drift_diff_const}

@@

\section{Drift and diffusion in a changing environment} \label{sec:drift_diff_chang}
\paragraph{Proving drift and diffusion terms in a changing environment.}
Let $x$ be the initial frequency of the invading phenotype and $X_t$ is the number of individuals with the phenotype at time $t$.
Then,
\begin{equation*}
E[X_t/N - x] \simeq x(1-x) S_t / N_e \;, 
\quad
\text{and}
\quad
V(X_t/N-x) \simeq  t x(1-x) / N_e \;,
\end{equation*}
where $S_t=\sum\limits_{i=1}^{t} N (1-\beta_t)$.

@@

Using \cref{eq:const_var} we get
\begin{equation}
V\left(\frac{X_{t+1}}{N} \bigg|X_t \right) 
= \frac{1}{N_e}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \;,
\end{equation}
and using the equation $y'(1-y') \simeq y(1-y)$ on the first part of \cref{eq:ch_var} we get
\begin{equation}\label{eq:ch_var2}
E\left[V\left(\frac{X_{t+1}}{N} \bigg|X_t \right)\right] 
= \frac{1}{N_e}E\left[\frac{X_t}{N}\left(1- \frac{X_t}{N}\right) \right] \simeq \frac{1}{N_e} x(1-x) \;.
\end{equation}
Moving on to simplify the second part of \cref{eq:ch_var} using \cref{eq:ch_var1},
\begin{equation}
V\left(E\left[\frac{X_{t+1}}{N} \bigg|X_t \right]\right) 
= V\left(\frac{X_t}{N} + \frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \;.\right) \right)
\end{equation}
Now, because $\frac{X_t}{N}$ is a frequency, i.e $0 \leq X_t/N \leq 1$, we know that $V\left(\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \right)\leq\frac{1}{4}$. We therefore find that
\begin{equation}
V\left(\frac{1}{N}s_{t+1}\frac{X_t}{N}\left(1-\frac{X_t}{N} \right) \right)
\leq \frac{1}{4N^2}s^2_{t+1} ;,
\end{equation}
and so it can be ignored.
Combining our equations we get
\begin{equation}
V\left(E\left[\frac{X_{t+1}}{N} \bigg|X_t \right]\right) 
= V\left(\frac{X_t}{N}\right) + O\left(\frac{1}{N^2}\right)\simeq V\left(\frac{X_t}{N}\right) \;.
\end{equation}
Using the induction assumption and \cref{eq:ch_var2},
\begin{equation}
V\left(\frac{X_{t+1}}{N}\right) 
\simeq \frac{1}{N_e}x(1-x) + \frac{1}{N_e}tx(1-x) \simeq \frac{1}{N_e}x(1-x)(t+1) \,
\end{equation}
which proves the diffusion term.
\end{proof}
\end{appendices}

%%%
